---
title: "tut01_coding"
author: "STA314 teaching team"
date: "2024-09-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tutorial 1: Linear Regression

```{r, message = FALSE}
rm(list=ls())# remove all variables in memory
## install.packages("...")
library(MASS)
library(ISLR)
```

For installation on Quarto, please consult: <https://quarto.org/docs/get-started/>

## Simple Linear Regression

### Basic intro with Boston housing dataset

Data regarding city of Boston: use `?Boston` for more information:

```{r}
# ?Boston
names(Boston) # what columns (predictors) we have
```

-   `medv` is median value of owner-occupied homes in thousands

-   `lstat` is lower status of the population

```{r}
#lm.fit=lm(medv~lstat) 
```

This gives us an error because we did not specify the data set #need to specify data set, hover over function to see arguments, to solve it:

1.  we can specify the data set we are using:

```{r}
lm.fit=lm(medv~lstat,data=Boston) 
```

2.  Or we could attach the data set:

```{r}
attach(Boston)
lm.fit=lm(medv~lstat)
# note that if the data is attached, there is no issue using the lm to fit the model directly.
```

We could take a quick look at the fitted model:

```{r}
lm.fit
```

Or a slightly more detailed summary:

```{r}
summary(lm.fit)
```

Note: Make sure you are able to interpret the results here, what does each value tell you?

To check specifically on coefficients, there are two ways that returns the same output:

```{r}
coef(lm.fit)
```

```{r}
lm.fit$coefficients
```

To access the fitted values of the model (the values of the dependent variable predicted by the model):

```{r}
# lm.fit$fitted.values this is a bit too long...
head(lm.fit$fitted.values) # this shows the first a few elements of the fitted value list
```

For more information on the input and output of the model, use:

```{r}
?lm  
```

### Confidence Interval with the fitted model

1.  we can check the confidence interval for regression coefficient:

```{r}
confint(lm.fit)
```

2.  the confidence interval for fitted values (the values of the dependent variable predicted by the model):

```{r}
predict(lm.fit,data.frame(lstat=(c(5,10,15))), interval="confidence")
```

3.  Similarly, the prediction interval for Y:

```{r}
predict(lm.fit,data.frame(lstat=(c(5,10,15))), interval="prediction")
```

### Graphical Interpretation of the linear model:

```{r}
plot(lstat,medv)
abline(lm.fit) #add our estimated model
abline(lm.fit,lwd=3)  #lwd represents the line width
abline(lm.fit,lwd=3,col="red")
```

**Discussion:** what do you notice about the data and/or the fit?

Meanwhile, we could also change the design of the plot:

```{r}
plot(lstat,medv,col="red")
plot(lstat,medv,pch=20) #plot using different symbols
plot(lstat,medv,pch="+")
plot(1:20,1:20,pch=1:20) #what the different sympols are in pch

```

Additionally, for multiple plots in one graph:

```{r}
par(mfrow=c(2,2)) # create 2 by 2 plot
plot(lm.fit) #very detailed figures for model fit
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))  #studentized residuals
```

## Multi-variate Linear Regression:

Using the same Boston Housing dataset, let's include another variable `age` into the equation:

```{r}
lm.fit=lm(medv~lstat+age,data=Boston)
summary(lm.fit)
```

We could also include all the variables in the data:

```{r}
lm.fit=lm(medv~.,data=Boston) #all covariates
summary(lm.fit)
```

What if we want to exclude some of the variables, for example excluding `age`?

```{r}
lm.fit1=lm(medv~.-age,data=Boston) #use all predictors except age
summary(lm.fit1)
```

### Interaction Terms

Please review the interpretation of coefficients when using interaction terms from the slides:

```{r}
#automatically include the main effect lstat and age
summary(lm(medv~lstat*age,data=Boston))

#an alternative approach
summary(lm(medv~lstat+age+lstat:age,data=Boston)) 
```

### Qualitative Predictors

Note the change now that we have non-continuous predictors, what do these represent?

Check out the `Carseats` data set in ISLR: <https://rdrr.io/cran/ISLR/man/Carseats.html>

```{r}
# ?Carseats
names(Carseats)
```

```{r}
summary(Carseats)
```

```{r}
lm.fit=lm(Sales~.+Income:Advertising+Price:Age,data=Carseats)
summary(lm.fit)
```

## Optional: explain usage of contrasts()

Feel free to check out the usage at:

<https://cran.r-project.org/web/packages/contrast/vignettes/contrast.html>
